\Large Clustering Trips Identified from Automatic Number \\ Plate Recognition Camera Scans

### 2018-02-21 classification.tex
Changed structure of the paper. Added main Methodology section with subsections
A few corrections and changes. Changed the name of length to degree, removed most references to future work. Rephrased a few sentences.

% !TEX root = paper.tex
\section{Clustering vehicles}\label{s.classification}

@@

Due to distinctly different traffic behaviour during weekends, we chose to consider only trips occurring during weekdays. We thus used all number-plate data collected between the 6th and 24th of February and exclude data from the 2 weekends in-between. Furthermore, as mentioned in section~\ref{s.trips}, we used fixed empirical values for $\Tau$. Table~\ref{t:trips-tau} displays the number of trips, average trip length and the proportion of trips of length one, for varying values of $\Tau$. The effect of incrementing $\Tau$ on resulting trips is clear: increased trip lengths and decreased trips containing just a single sighting. On the other hand, we did not set a value for $\tau$, but instead handled implausible trips by filtering all sightings with confidence below 85\%. Duplicates were filtered after identifying consecutive sightings by the same camera occurring within the same trip. Clock synchronisation errors were provided in milliseconds with none exceeding 5 seconds. These were therefore ignored.

\begin{table}[t]
\centering
\tabcolsep=0.25cm
\begin{tabular}{c c c c}
  \hline
Tau & Trips & \thead{Average\\Length} & \thead{Proportion of Trips \\ Length 1} \\
  \hline
5 min & 13,603,759 & 1.46 & 0.70 \\
7.5 min & 12,394,709 & 1.60 & 0.64 \\
10 min & 11,690,791 & 1.69 & 0.61 \\
15 min & 10,823,333 & 1.83 & 0.57 \\
20 min & 10,305,860 & 1.92 & 0.54 \\
30 min &  9,653,499 & 2.04 & 0.52 \\
   \hline
\end{tabular}
\caption{Overview of trip data for varying values of $\Tau$.}
\label{t:trips-tau}
\vspace{-0.5cm}
\end{table}

Transforming trips into features which can be used in clustering algorithms was a 3-step process:
\begin{enumerate*}[label=(\roman*)]
  \item First, every trip was summarised as a single row of data. The following information was extracted: length of trip, origin, destination, route, start and end times.
  \item Second, daily trip information was obtained for each vehicle: number of trips, median of trip lengths, number of sightings, distinct number of origins destinations, and routes, hour of first sighting, hour of last sighting and total rest time between trips.
  \item Finally, daily information per vehicle was collapsed into a single row by averaging this information across the 15 days.
\end{enumerate*}

Table~\ref{t:features} depicts a sample of the resulting features vector. A total of 1034107 distinct vehicles were detected. However, because there is a high percentage of trips containing a single sighting, some of these features were highly correlated. We therefore, chose to remove three of the features represented in~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}. Furthermore we considered that a trip of length one has no destination (which explains values of average distinct destination below zero) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642006 unique vehicles.

\begin{table*}[t]
\centering
\begin{tabular}{c c c c c c c c c c c}
  \hline
 \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Length} & \thead{Average\\Sightings} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\Distinct\\Routes} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Hour\\Difference} & \thead{Average\\Rest\\Time} \\
  \hline
41 & 3.42 & 1.25 & 5.75 & 2.75 & 1.00 & 3.00 & 15.33 & 19.25 & 3.80 & 3.70 \\
3 & 1.50 & 1.00 & 1.50 & 1.50 & 0.00 & 1.50 & 14.00 & 14.50 & 0.73 & 0.73 \\
7 & 2.33 & 1.33 & 3.33 & 2.33 & 0.67 & 2.33 & 11.00 & 13.00 & 2.44 & 2.41 \\
12 & 2.40 & 1.10 & 3.60 & 2.40 & 0.60 & 2.40 & 14.40 & 16.40 & 2.00 & 1.95 \\
   \hline
\end{tabular}
\caption{Sample of extracted features from trips taken from 15 weekdays of number plate data.}
\vspace{-0.2cm}
\label{t:features}
\end{table*}

@@

Table~\ref{t:tau_comparison} provides a summary of multiple runs of \emph{k-means}. For each value of $\Tau$, the optimal number of clusters is selected by picking the value of $k$ that maximises the Calinski-Harabasz criterion. Furthermore, the measures of inter-cluster (betweeness) and intra-cluster (withiness) are depicted relative to the corresponding total sum of squares (total betweeness and total withiness). It is noteworthy that the best value of $k$ increases inversely to $\Tau$. Although a higher value of $k$ seems to suggest that trips with smaller values of $\Tau$ are able capture the variance in the data better, we have to consider that varying $\Tau$ affects the average trip length in the same direction whilst affecting the total number of trips in the opposite direction (table~\ref{t:trips-tau}).

@@

\begin{table}[t]
\centering
\tabcolsep=0.17cm
\begin{tabular}{c c c c c}
  \hline
Tau & Best $k$ & Betweenss & \thead{Average\\Withinss} & \thead{Calinski-\\Harabasz} \\
  \hline
5 min &   8 & 0.923 & 0.125 & 1,116,962 \\
  7.5 min &   7 & 0.904 & 0.143 & 1,003,744 \\
  10 min&   7 & 0.896 & 0.143 &   911,044 \\
  15 min &   6 & 0.865 & 0.167 &   794,557 \\
  20 min &   4 & 0.786 & 0.250 &   754,241 \\
  30 min &   3 & 0.710 & 0.333 &   743,001 \\
   \hline
\end{tabular}
\caption{\emph{k-means} performance for several values of $\Tau$.}
\label{t:tau_comparison}
\vspace{-0.5cm}
\end{table}

\begin{table*}[t]
\centering
\small
\begin{tabular}{c c c c c c c c c c}
  \hline
 Cluster &  Size & \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Length} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Rest\\Time} \\
  \hline
  1 & 108868 & 30.67 & 2.91 & 1.56 & 2.65 & 1.09 & 9.67 & 16.11 & 6.35 \\
  2 & 2948 & 170.08 & 13.69 & 1.47 & 9.20 & 4.67 & 6.89 & 18.69 & 11.65 \\
  3 & 309748 & 5.95 & 2.01 & 1.44 & 1.92 & 0.68 & 12.52 & 14.50 & 1.93 \\
  4 & 163982 & 16.99 & 2.30 & 1.46 & 2.13 & 0.77 & 11.12 & 14.99 & 3.81 \\
  5 & 45414 & 50.30 & 4.26 & 1.52 & 3.67 & 1.57 & 9.04 & 16.86 & 7.69 \\
  6 & 10380 & 86.99 & 7.16 & 1.43 & 5.55 & 2.49 & 8.41 & 17.60 & 9.03 \\
  7 & 666 & 331.50 & 26.43 & 1.57 & 9.31 & 5.17 & 5.73 & 19.59 & 13.14 \\
   \hline
\end{tabular}
\caption{Clusters sizes and centres for $\Tau = 7.5$ minutes.}
\label{t:kmeans_centers_450}
\end{table*}

\begin{table*}[t]
\centering
\small
\begin{tabular}{c c c c c c c c c c}
  \hline
 Cluster &  Size & \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Length} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Rest\\Time} \\
  \hline
  1 & 371318 & 7.19 & 1.76 & 1.68 & 1.68 & 0.71 & 12.30 & 14.55 & 2.14 \\
    2 & 49852 & 44.85 & 3.73 & 1.90 & 3.16 & 1.73 & 8.98 & 17.02 & 7.76 \\
    3 & 189224 & 22.82 & 2.29 & 1.84 & 2.11 & 1.06 & 10.00 & 15.76 & 5.60 \\
    4 & 4593 & 114.34 & 9.12 & 2.30 & 5.92 & 3.98 & 6.93 & 18.21 & 10.43 \\
   \hline
\end{tabular}
\caption{Clusters sizes and centres for $\Tau = 20$ minutes.}
\label{t:kmeans_centers_1200}
\end{table*}

>>>>

% !TEX root = paper.tex
\subsection{Clustering trips}\label{s.classification}

@@

Due to distinctly different traffic behaviour during weekends, we chose to consider only trips occurring during weekdays. We thus used all number-plate data collected between the 6th and 24th of February and exclude data from the 2 weekends in-between. Furthermore, as mentioned in section~\ref{s.trips}, we used fixed empirical values for $\Tau$. Table~\ref{t:trips-tau} displays the number of trips, average trip degree and the proportion of trips of degree one, for varying values of $\Tau$. The effect of incrementing $\Tau$ on resulting trips is clear: increased trip degrees and decreased trips containing just a single sighting. On the other hand, we did not set a value for $\tau$, but instead handled implausible trips by filtering all sightings with confidence below 85\%. Duplicates were filtered after identifying consecutive sightings by the same camera occurring within the same trip. Clock synchronisation errors were provided in milliseconds with none exceeding 5 seconds. These were therefore ignored.

\begin{table}[t]
\centering
\tabcolsep=0.25cm
\begin{tabular}{c c c c}
  \hline
Tau & Trips & \thead{Average\\Degree} & \thead{Proportion of Trips \\ Degree 1} \\
  \hline
5 min & 13,603,759 & 1.46 & 0.70 \\
7.5 min & 12,394,709 & 1.60 & 0.64 \\
10 min & 11,690,791 & 1.69 & 0.61 \\
15 min & 10,823,333 & 1.83 & 0.57 \\
20 min & 10,305,860 & 1.92 & 0.54 \\
30 min &  9,653,499 & 2.04 & 0.52 \\
   \hline
\end{tabular}
\caption{Overview of trip data for varying values of $\Tau$.}
\label{t:trips-tau}
\vspace{-0.5cm}
\end{table}

Transforming trips into features which can be used in clustering algorithms was a 3-step process:
\begin{enumerate*}[label=(\roman*)]
  \item First, every trip was summarised as a single row of data. The following information was extracted: degree of trip, origin, destination, route, start and end times.
  \item Second, daily trip information was obtained for each vehicle: number of trips, median of trip degrees, number of sightings, distinct number of origins destinations, and routes, hour of first sighting, hour of last sighting and total rest time between trips.
  \item Finally, daily information per vehicle was collapsed into a single row by averaging this information across the 15 days.
\end{enumerate*}

Table~\ref{t:features} depicts a sample of the resulting features vector. A total of 1034107 distinct vehicles were detected. However, because there is a high percentage of trips containing a single sighting, some of these features were highly correlated. We therefore, chose to remove three of the features represented in~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}. Furthermore we considered that a trip of degree one has no destination (which explains values of average distinct destination below zero) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642006 unique vehicles.

\begin{table*}[t]
\centering
\begin{tabular}{c c c c c c c c c c c}
  \hline
 \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Degree} & \thead{Average\\Sightings} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\Distinct\\Routes} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Hour\\Difference} & \thead{Average\\Rest\\Time} \\
  \hline
41 & 3.42 & 1.25 & 5.75 & 2.75 & 1.00 & 3.00 & 15.33 & 19.25 & 3.80 & 3.70 \\
3 & 1.50 & 1.00 & 1.50 & 1.50 & 0.00 & 1.50 & 14.00 & 14.50 & 0.73 & 0.73 \\
7 & 2.33 & 1.33 & 3.33 & 2.33 & 0.67 & 2.33 & 11.00 & 13.00 & 2.44 & 2.41 \\
12 & 2.40 & 1.10 & 3.60 & 2.40 & 0.60 & 2.40 & 14.40 & 16.40 & 2.00 & 1.95 \\
   \hline
\end{tabular}
\caption{Sample of extracted features from trips taken from 15 weekdays of number plate data.}
\vspace{-0.2cm}
\label{t:features}
\end{table*}

@@

Table~\ref{t:tau_comparison} provides a summary of multiple runs of \emph{k-means}. For each value of $\Tau$, the optimal number of clusters is selected by picking the value of $k$ that maximises the Calinski-Harabasz criterion. Furthermore, the measures of inter-cluster (betweeness) and intra-cluster (withiness) are depicted relative to the corresponding total sum of squares (total betweeness and total withiness). It is noteworthy that the best value of $k$ increases inversely to $\Tau$. Although a higher value of $k$ seems to suggest that trips with smaller values of $\Tau$ are able capture the variance in the data better, we have to consider that varying $\Tau$ affects the average trip degree in the same direction whilst affecting the total number of trips in the opposite direction (table~\ref{t:trips-tau}).

@@

\begin{table}[t]
\centering
\tabcolsep=0.17cm
\begin{tabular}{c c c c c}
  \hline
Tau & Best $k$ & Betweenss & \thead{Average\\Withinss} & \thead{Calinski-\\Harabasz} \\
  \hline
5 min &   8 & 0.923 & 0.125 & 1,116,962 \\
  7.5 min &   7 & 0.904 & 0.143 & 1,003,744 \\
  10 min&   7 & 0.896 & 0.143 &   911,044 \\
  15 min &   6 & 0.865 & 0.167 &   794,557 \\
  20 min &   4 & 0.786 & 0.250 &   754,241 \\
  30 min &   3 & 0.710 & 0.333 &   743,001 \\
   \hline
\end{tabular}
\caption{\emph{k-means} performance for several values of $\Tau$.}
\label{t:tau_comparison}
\end{table}

\begin{table}[t]
\centering
\tabcolsep=0.15cm
\begin{tabular}{c c c c c c}
  \hline
 Cluster & Size & \thead{Average\\Trips} & \thead{Average\\Distinct\\Origins} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} \\
  \hline
1 & 157962 & 2.45 & 2.27 & 11.38 & 14.93 \\
  2 & 303513 & 2.18 & 2.09 & 12.53 & 14.47 \\
  3 & 21549 & 6.02 & 5.00 & 8.68 & 17.23 \\
  4 & 108094 & 2.99 & 2.73 & 9.99 & 15.86 \\
  5 & 509 & 33.62 & 10.42 & 5.81 & 19.67 \\
  6 & 1993 & 17.99 & 11.24 & 6.45 & 19.09 \\
  7 & 4971 & 10.41 & 7.64 & 7.97 & 18.04 \\
  8 & 58059 & 4.09 & 3.62 & 9.22 & 16.60 \\
   \hline
\end{tabular}
\caption{Cluster sizes and centres for $\Tau = 5$ minutes.}
\label{t:kmeans_centers_300}
\end{table}

### 2018-02-21 dataset.tex
Final version finished. Finished results description and fixed bad boxes in bibliography. Finally managed to fit everything with a few vspaces here and there
Changed structure of the paper. Added main Methodology section with subsections
A few corrections and changes. Changed the name of length to degree, removed most references to future work. Rephrased a few sentences.

% !TEX root = paper.tex
\section{Tyne and Wear ANPR Data}\label{s.ncl}

Automatic number-plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems. They can be used by government subsidised entities in urban traffic management and control; by commissioned highway agencies in electronic toll collection; or by law enforcement organisations in detecting speeding vehicles and validating number plate registrations. The wide diversity of applications, paired with the large improvements in price-to-performance ratios of ANPR hardware and software systems, has resulted in increased investments of ANPR cameras for urban environments~\cite{EvolutionUTMC2013, SurveyITS2011}.

\begin{figure*}[t]
\centering
\begin{subfigure}[t]{.48\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{observations_per_day.pdf}
  \caption{Total number of scans recorded per day in Tyne and Wear. There is a clear seasonal effect caused by decreasing traffic demands at weekends and increasing traffic volume during weekdays.}
  \label{fig:observations-per-day}
\end{subfigure}\hfill
\begin{subfigure}[t]{.48\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{observations_per_day_camera.pdf}
  \caption{Number of scans recorded per ANPR camera and day in Tyne and Wear. Inter-camera variability is observed, as some cameras are located in more traffic intensive road sections than others. Decommissioned or temporarily unavailable cameras (due to loss of power, faulty camera, road closed, etc) are depicted at the bottom of the graph.}
  \label{fig:observations-per-camera-day}
\end{subfigure}
\caption{License plate scans recorded by ANPR cameras during February 2017, in the region of Tyne and Wear, United Kingdom.}
\label{fig:time-series}
\end{figure*}

@@

Number plate data, in its essence, is a stream of events, each representing a vehicle observed by one camera at a specific point in time. An excerpt of the data can be found in table~\ref{t:np_data_example}. All number-plates were anonymised by the UTMC through a hashing algorithm before the data was shared. Cameras are uniquely identified by an integer and timestamps are relative to each camera's clock. However, the cameras are connected through a private network which provides clock synchronisation using the Network Time Protocol (NTP). Therefore, the timestamps can be used directly if the synchronisation error is negligible. The following additional information is also captured and provided by each camera:
\begin{enumerate*}[label=(\roman*)]
  \item the clock synchronisation error (milliseconds);
  \item the camera's confidence that the identified number plate is the true number plate (percentage);
  \item the direction of travel, away or towards the camera.
\end{enumerate*}
The confidence in the observation is especially useful as it helps diagnosing license plate recognition errors. On the other hand, the direction of travel is dependent upon the orientation of the camera, which is not provided. Hence, we chose to ignore the latter in this work, and aim to explore this information in future works.

\begin{table}[ht]
\centering
\tabcolsep=0.17cm
\small
\begin{tabular}{c c c c c}
  \hline
  Vehicle & Camera & Timestamp & \thead{Clock \\Error} & Confidence \\
  \hline
  169239 & 1031 & 1454284800.26 &   0 & 100 \\
  12862943 & 18 & 1454284800.97 &   8 &  61 \\
  16243894 & 22 & 1454284801.46 &   6 &  86 \\
  4817789 & 52 & 1454284803.43 &  13 &  94 \\
  5503486 & 110 & 1454284802.19 &  22 &  91 \\
  15244177 & 115 & 1454284802.83 &  18 &  87 \\
   \hline
\end{tabular}
\caption{Sample of number plate data. Clock error is given in milliseconds and confidence as a percentile value.}
\label{t:np_data_example}
%\vspace{-2cm}
\end{table}

>>>>

% !TEX root = paper.tex
\section{Methodology}
\subsection{Tyne and Wear ANPR Data}\label{s.ncl}

\begin{table}[!ht]
\centering
\tabcolsep=0.17cm
\small
\begin{tabular}{c c c c c}
  \hline
  Vehicle & Camera & Timestamp & \thead{Clock \\Error} & Confidence \\
  \hline
  169239 & 1031 & 1454284800.26 &   0 & 100 \\
  12862943 & 18 & 1454284800.97 &   8 &  61 \\
  16243894 & 22 & 1454284801.46 &   6 &  86 \\
  4817789 & 52 & 1454284803.43 &  13 &  94 \\
  5503486 & 110 & 1454284802.19 &  22 &  91 \\
  15244177 & 115 & 1454284802.83 &  18 &  87 \\
   \hline
\end{tabular}
\caption{Sample of number plate data. Clock error is given in milliseconds and confidence as a percentile value.}
\label{t:np_data_example}
%\vspace{-2cm}
\end{table}

Automatic num\-ber plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems. They can be used by government subsidised entities in urban traffic management and control; by commissioned highway agencies in electronic toll collection; or by law enforcement organisations in detecting speeding vehicles and validating number plate registrations. The wide diversity of applications, paired with the large improvements in price-to-performance ratios of ANPR hardware and software systems, has resulted in increased investments of ANPR cameras for urban environments~\cite{EvolutionUTMC2013, SurveyITS2011}.

\begin{figure*}[t]
\centering
\begin{subfigure}[t]{.48\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{observations_per_day.pdf}
  \caption{Total number of scans recorded per day in Tyne and Wear. There is a clear seasonal effect caused by decreasing traffic demands at weekends and increasing traffic volume during weekdays.}
  \label{fig:observations-per-day}
\end{subfigure}\hfill
\begin{subfigure}[t]{.48\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{observations_per_day_camera.pdf}
  \caption{Number of scans recorded per ANPR camera and day in Tyne and Wear. Inter-camera variability is observed, as some cameras are located in more traffic intensive road sections than others. Decommissioned or temporarily unavailable cameras (due to loss of power, faulty camera, road closed, etc) can be identified at the bottom.}
  \label{fig:observations-per-camera-day}
\end{subfigure}
\caption{License plate scans recorded by ANPR cameras during February 2017, in the region of Tyne and Wear, United Kingdom.}
\label{fig:time-series}
\end{figure*}

@@

Number plate data, in its essence, is a stream of events, each representing a vehicle observed by one camera at a specific point in time. An excerpt of the data can be found in table~\ref{t:np_data_example}. All number plate were anonymised by the UTMC through a hashing algorithm before the data was shared. Cameras are uniquely identified by an integer and timestamps are relative to each camera's clock. Clock synchronisation is performed using the Network Time Protocol (NTP), as the cameras are connected, through a private network, to a central server. Therefore, the recorded timestamps can be used directly if the synchronisation error is negligible. The following additional information is also captured and provided by each camera:
\begin{enumerate*}[label=(\roman*)]
  \item the clock synchronisation error (milliseconds);
  \item the camera's confidence that the identified number plate is the true number plate (percentage);
  \item the direction of travel, away or towards the camera.
\end{enumerate*}
The confidence in the observation is especially useful as it helps diagnosing license plate recognition errors. On the other hand, the direction of travel is dependent upon the orientation of the camera, which is not provided. Hence, we chose to ignore the latter in this work.

### 2018-02-21 discussion.tex
A few corrections and changes. Changed the name of length to degree, removed most references to future work. Rephrased a few sentences.

Most urban cities in the world possess a network of ANPR cameras that are used for law enforcement as well as traffic monitoring and control. Number plate data collected in the Tyne and Wear area is stored and leveraged by the UTMC for computing average journey times across a selection of sensitive roads. However, number plate data could be used more extensively to identify and study individual and collective travel patterns. In this paper, we have presented a set of definitions and constraints that establish a conceptual foundation for identifying vehicle trips from number plate detections. We also identify two parameters, $\tau$ and $\Tau$ as critical in the discrimination of plausible and implausible trips. Hence, future work should first and foremost focus on developing formal methods to estimate these parameters from observed distributions of travel times and by applying knowledge about the structure of the road network. Moreover, methods for addressing issues concerning camera performance, namely wrong and duplicate scans, should be further developed and researched.

@@

>>>>

Most urban cities in the world employ a network of ANPR cameras that are used for law enforcement as well as traffic monitoring and control. Number plate data collected in the Tyne and Wear area is stored and leveraged by the UTMC for computing average journey times across a selection of sensitive roads. However, number plate data could be used more extensively to identify and study individual and collective travel patterns. In this paper, we have presented a set of definitions and constraints that establish a conceptual foundation for identifying vehicle trips from number plate detections. We also identify two parameters, $\tau$ and $\Tau$ as critical in the discrimination of plausible and implausible trips. Hence, future work should first and foremost focus on developing formal methods to estimate these parameters from observed distributions of travel times and by applying knowledge about the structure of the road network. Moreover, methods for addressing issues concerning camera performance, namely wrong and duplicate scans, should be further developed and researched.

@@

Hence, we chose to ignore camera direction in this work, and aim to explore this information in future works.

On the other hand, a consistent set of models and methodologies for estimating $\Tau$ and $\tau$, under a variety of circumstances and road anomalies, remains an open research problem. Nevertheless, the importance of estimating these parameters should not be understated. Any errors in trip identification that occur due to poor estimation and filtering methods will propagate and will be amplified in posterior analysis done using trip data. Therefore, $\Tau$ and $\tau$ estimation is something that needs to be carefully addressed and studied in the future in order to fully understand its impact in the outcome of the research.

### 2018-02-21 intro.tex
Fixed the following issues: Font color red to black, formula 4.3 fixed, re-written paragraph about constraint 4.4
A few corrections and changes. Changed the name of length to degree, removed most references to future work. Rephrased a few sentences.

In a utopian model we would have full disclosure of all journeys made by all vehicles on the road infrastructure. However, this has numerous ethical and technical issues. From an ethical standpoint should we be allowed to know where all vehicles are at any given point in time. From a technical point of view, although every vehicle could be fitted with a GPS tracker -- costly in its own right -- there would still exist the issue of how we would collect and stream all of this data for future processing. Alternatively one can view the problem the other way around and rather than tracking individual vehicles look at collecting information by observing vehicles passing points within the road networks. A prime example of this approach are Automatic Number-plate Recognition (ANPR) cameras. These cameras are a combination of digital camera coupled with Artificial Intelligence to identify number-plates within the image and convert these into strings of characters. ANPR cameras are normally fixed in location\footnote{Although cameras can be in a vehicle and moved from location to location.} able to view all vehicles passing that location.

For ANPR the problem now becomes that of recovering as much information about vehicle's journey as possible from the limited number of observations. ANPR cameras are normally located on major roads and interchanges, however, this only covers a tiny fraction of the road network. We can, though, estimate routes between cameras by understanding the distances between cameras and the most ``sensible'' routes between them. This allows us, given a set of ANPR sightings of the same vehicle, to produce a ``most likely'' route for that journey. It should be noted that we cannot determine the actual start and end of the journey as these will happen in areas not covered by ANPR. It should also be noted that for ethical reasons it is not normal to obtain actual number-plates, but rather the hash of these. Though, for most situations this will suffice.

@@

\begin{itemize}
	\item {\bf Number-plate miss-reads:} Although ANPR cameras have accuracies of around {\color{red}99\%}, miss-reads are possible. This can lead to sightings being missed or vehicles being wrongly sighted in locations.
	\item {\bf Timing errors:} The time-stamps of sightings could be erroneous. The minor side of this is implausible journey times, though, more seriously, this can lead to reordering the set of cameras on a particular journey.
	\item {\bf Clones number-plates:} For various reasons a number-plate may be cloned and used on a different vehicle. This can lead to impossible journeys and journeys that the real vehicle did not make.
\end{itemize}

@@

The main use of ANPR data for the UTMC is estimating average journey times for selected or sensitive links in the road network. Furthermore, several authors have extensively researched how to use number-plate data as an extension to link counts for estimating origin-destination matrices and link flows~\cite{Castillo2010, Castillo2008, Hazelton2012}. However, very few works have focused on analysing individual or collective travel patterns from number-plate data, particularly across extended periods of time. Moreover, there is no consistent conceptual and analytical framework for transforming number plate data into a historical sequence of trips for each vehicle. Finally, we believe that trip data, properly identified from number-plate data, has the potential to unlock a number of new applications for urban traffic control and law enforcement. Thus, in section~\ref{s.trips} we present a conceptual methodology for grouping multiple camera observations of the same vehicle into one or several trips of that vehicle.

Determining the distribution of travel modes is one of the fundamental steps in the four-stage model: an essential traffic modelling methodology for transportation planning~\cite{FourStepModel}. Previous works have used trip information derived from different sources of data to identify travel mode or purpose of trip. More notably, survey data, floating car data and mobile phone data have been used~\cite{ODMobileData, ClusteringGPS}.  Although number plate data has been used in~\cite{Clustering}, to identify different categories of trips, the authors do not differentiate between private or public travel modes and focus instead on categorising trips by time of occurrence. Hence, in section~\ref{s.classification} we apply the \emph{k-means} clustering algorithm to derived trip data and based on the results, we discuss the limitations of proposed methods and that of ANPR data.

>>>>

In a utopian model we would have full disclosure of all journeys made by all vehicles on the road infrastructure. However, this has numerous ethical and technical issues. From an ethical standpoint should we be allowed to know where all vehicles are at any given point in time. From a technical point of view, although every vehicle could be fitted with a GPS tracker -- costly in its own right -- there would still exist the issue of how we would collect and stream all of this data for future processing. Alternatively one can view the problem the other way around and rather than tracking individual vehicles look at collecting information by observing vehicles passing points within the road networks. A prime example of this approach are Automatic Number Plate Recognition (ANPR) cameras. These cameras are a combination of digital camera coupled with Artificial Intelligence to identify number plates within the image and convert these into strings of characters. ANPR cameras are normally fixed in location\footnote{Although cameras can be in a vehicle and moved from location to location.} able to view all vehicles passing that location.

For ANPR the problem now becomes that of recovering as much information about vehicle's journey as possible from the limited number of observations. ANPR cameras are normally located on major roads and interchanges, however, this only covers a tiny fraction of the road network. We can, though, estimate routes between cameras by understanding the distances between cameras and the most ``sensible'' routes between them. This allows us, given a set of ANPR sightings of the same vehicle, to produce a ``most likely'' route for that journey. It should be noted that we cannot determine the actual start and end of the journey as these will happen in areas not covered by ANPR. It should also be noted that for ethical reasons it is not normal to obtain actual number plates, but rather the hash of these. Though, for most situations this will suffice.

@@

\begin{itemize}
	\item {\bf Number plate miss-reads:} Although ANPR cameras have accuracies of around 99, miss-reads are possible. This can lead to sightings being missed or vehicles being wrongly sighted in locations.
	\item {\bf Timing errors:} The time-stamps of sightings could be erroneous. The minor side of this is implausible journey times, though, more seriously, this can lead to reordering the set of cameras on a particular journey.
	\item {\bf Clones number plates:} For various reasons a number plate may be cloned and used on a different vehicle. This can lead to impossible journeys and journeys that the real vehicle did not make.
\end{itemize}

@@

The main use of ANPR data for the UTMC is estimating average journey times for selected or sensitive links in the road network. Furthermore, several authors have extensively researched how to use number plate data as an extension to link counts for estimating origin-destination matrices and link flows~\cite{Castillo2010, Castillo2008, Hazelton2012}. However, very few works have focused on analysing individual or collective travel patterns from number plate data, particularly across extended periods of time. Moreover, there is no consistent conceptual and analytical framework for transforming number plate data into a historical sequence of trips for each vehicle. Finally, we believe that trip data, properly identified from number plate data, has the potential to unlock a number of new applications for urban traffic control and law enforcement. Thus, in section~\ref{s.trips} we present a conceptual methodology for grouping multiple camera observations of the same vehicle into one or several trips of that vehicle.

Determining the distribution of travel modes is one of the fundamental steps in the four-stage model: an essential traffic modelling methodology for transportation planning~\cite{FourStepModel}. Previous works have used trip information derived from different sources of data to identify travel mode or purpose of trip. More notably, survey data, floating car data and mobile phone data have been used~\cite{ODMobileData, ClusteringGPS}.  Although number plate data has been used in~\cite{Clustering} to identify different categories of trips, the authors do not differentiate between private or public travel modes and focus instead on categorising trips by time of occurrence. Hence, in section~\ref{s.classification} we apply the \emph{k-means} clustering algorithm to derived trip data and based on the results, we discuss the limitations of proposed methods and that of ANPR data.
